---
title: "RepData_PeerAssignment2"
author: "thecapacity"
date: "January 12, 2015"
output: html_document
---

## Title: NOAA Storm Analysis

**The Results were as follows... ##```"FILL THIS IN"```**

### Synopsis:

**```THIS should be a synopsis which describes and summarizes analysis in at most 10 complete sentences.```**

#### Background

This the R Markdown document associated with my repository located at: https://github.com/thecapacity/RepData_PeerAssessment2

This document will be used to capture the results of my data analysis in order to make them reproducable, and will be published at my [RPubs Account](http://rpubs.com/thecapacity/). This document will represent a stand alone assessment, but for more details please check out the [GitHub Repository](http://rpubs.com/thecapacity/).

```{r setoptions}
### Setup some defaults

# Ensure pristine working environment
## rm(list=ls())
## This has been commented out for submission to ensure no disruptivie side-effets for others

library(knitr)
opts_chunk$set(echo = TRUE, fig.path="figure/", dev="png")
```

These global defaults are set, or suggested ```# As comments``` to promote consistent behavior.

This work was done on a Macbook, running OSX 10.9 with the software stack summarized as follows:
```{r}
# Summarize the Operating Environment

search()
```

**This analysis also assumes the bzunzip2 command is available to extract the data via the command line.**

## Data Processing

This section outlines (in words and code) how the data were loaded into R and processed for subsequent analysis. **Analysis will start from the raw CSV file containing the data; and there will be no preprocessing outside of this document.**

**If preprocessing is time-consuming the cache = TRUE option may be used for certain code chunks.**

```{r dataDownload, cache=TRUE }
## Data Processing Code is here, to load and format the data
##      Subsequent analysis, actually deriving results is captured in the analysis sub-section.

data_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
data_file <- "data_dir/data"

if (! file.exists("data_dir")) {
    dir.create("data_dir")
    bz_data_file <- "./data_dir/data.bz2"
    download.file(data_url, mode="wb", destfile=bz_data_file, method="curl")
    dateDownloaded <- date()
    system2("bunzip2", args=c("-dfq", "data_dir/data.bz2") )
}

## NOTE THIS IS JUST 100 ROWS FOR TESTING PURPOSES!!!!
data_file <- "data_dir/test_data"
# Note, data is extracted every time - but not necessarially downloaded - to ensure consistency
if ( file.exists(data_file)) {
    my_data <- read.csv(data_file)
}
```

The data used for this analysis was downloadd on ```r dateDownloaded```.

The first two lines for this data file are:
```{r}
readLines(data_file, 2)
```

#### Analysis

**The document must have at least one figure containing a plot.**

**The must have no more than three figures. Figures may have multiple plots in them (i.e. panel plots), but there cannot be more than three figures total.**

```{r analysis }
## This captures the actual analysis and activities to derive final data
##      Visual charts and graphs are presented in the Results section

```

## Results

This Results section will present the final results.

**The document must have at least one figure containing a plot.**

**The document must have no more than three figures. Figures may have multiple plots in them (i.e. panel plots), but there cannot be more than three figures total.**

```{r resutls }
## Final Graphs and a summary of conclusions will be captured here, with
##      all work being done bt the Data Processing Section (and the Analysis Sub-Sectin)

```
